---
title: "Create and share a SeRiouS capsule"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create and share a SeRiouS capsule}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Create and share a SeRiouS capsule

SeRiouS is an R/Shiny framework for creating interactive learning capsules.

A SeRiouS capsule is organised as a pedagogical path made of successive steps. Each step may contain:

- an objective;
- a piece of R code;
- a question;
- a console output;
- a plot;
- a PDF or other resource.

The general idea is simple:

```text
SeRiouS provides the stage.
The teacher writes the play.
```

This vignette explains how to create, check, run and share a capsule as a simple folder.

# Create a capsule folder

The easiest way to start is to create a capsule skeleton.

```{r}
library(SeRiouS)

create_capsule_skeleton(
  path = "my_capsule",
  title = "My first SeRiouS capsule"
)
```

This creates a folder named `my_capsule`.

# Folder structure

A capsule folder has the following structure:

```text
my_capsule/
├── capsule.R
├── serious.yml
├── README.md
├── data/
├── pdf/
├── img/
└── www/
```

The most important file is:

```text
capsule.R
```

It must define a function named `create_capsule()`.

This function must return a `learning_capsule` object.

The other folders are optional resource folders:

```text
data/   datasets used by the capsule
pdf/    PDF supports or slides
img/    images used by the capsule
www/    web resources for the Shiny app
```

# Open and edit `capsule.R`

The generated `capsule.R` file contains a minimal working example.

A typical capsule is built in three steps:

```{r}
sections <- SeRiouS::make_sections(...)

steps <- list(
  SeRiouS::make_step(...),
  SeRiouS::make_step(...)
)

SeRiouS::build_linear_capsule(
  id = "my_capsule",
  title = "My capsule",
  method = "My method",
  steps = steps,
  sections = sections
)
```

The function `make_step()` defines one step of the learning path.

For example:

```{r}
SeRiouS::make_step(
  id = "intro",
  title = "Discover the data",
  section = "data",
  objective = "Display the first rows of the iris dataset.",
  code = "head(iris)",
  question = "Which dataset is used in this capsule?",
  expected_answer = "iris"
)
```

A step can also generate a plot:

```{r}
SeRiouS::make_step(
  id = "plot",
  title = "Create a plot",
  section = "visual",
  objective = "Plot sepal length against petal length.",
  outputs = c("console", "plot"),
  code = paste(
    "plot(",
    "  iris$Sepal.Length,",
    "  iris$Petal.Length,",
    "  pch = 16",
    ")",
    sep = "\n"
  ),
  question = "Which base R function creates this graph?",
  expected_answer = "plot"
)
```

# Check the capsule

Before sharing a capsule, it is recommended to check it.

From inside the capsule folder:

```{r}
check_capsule_dir(".")
```

Or from the parent folder:

```{r}
check_capsule_dir("my_capsule")
```

This function checks that:

```text
capsule.R exists
create_capsule() exists
create_capsule() returns a learning_capsule object
the capsule can be validated
the main folder structure is coherent
```

# Run the capsule

To run the capsule from the parent folder:

```{r}
run_capsule_dir("my_capsule")
```

To run it from inside the capsule folder:

```{r}
run_capsule_dir(".")
```

This launches the Shiny interface.

# Share the capsule

To share the capsule, zip the folder:

```text
my_capsule/
```

The recipient only needs to install SeRiouS and run:

```{r}
library(SeRiouS)

run_capsule_dir("my_capsule")
```

This makes it possible to share pedagogical material without creating a full R package.

# Summary

The typical workflow for a contributor is:

```{r}
library(SeRiouS)

create_capsule_skeleton("my_capsule")
check_capsule_dir("my_capsule")
run_capsule_dir("my_capsule")
```

Then edit:

```text
my_capsule/capsule.R
```

and share the folder when the capsule is ready.
