---
title: "Create and share a SeRiouS capsule"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create and share a SeRiouS capsule}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Create and share a SeRiouS capsule

SeRiouS is an R/Shiny framework for creating interactive learning capsules.

A SeRiouS capsule is organised as a pedagogical board made of **cells**
connected by **links**. Each cell may contain:

- a learning objective;
- explanatory text;
- a piece of R code;
- a question used to unlock progress;
- console output;
- plots;
- PDF files or other resources.

The general idea is simple:

```text
SeRiouS provides the stage.
The teacher writes the play.
```

This vignette explains how to create, inspect, edit, check, run and share a
capsule as a simple folder.

# The recommended workflow

The recommended way to work with SeRiouS is to create a capsule folder.

```{r}
library(SeRiouS)

create_capsule_skeleton(
  path = "my_capsule",
  title = "My first SeRiouS capsule"
)
```

This creates a folder named `my_capsule`.

A capsule folder is easy to share: it can be zipped, sent to another user, or
stored in a version-controlled repository.

# Folder structure

A SeRiouS capsule folder has the following structure:

```text
my_capsule/
├── serious.yml
├── README.md
├── cells/
├── data/
├── pdf/
├── img/
└── www/
```

The main elements are:

```text
serious.yml   global metadata, sections, colors, packages and start cell
cells/        one YAML file per pedagogical cell
data/         optional datasets
pdf/          optional PDF resources
img/          optional images
www/          optional web assets such as CSS
```

The most important folder is `cells/`. Each cell of the pedagogical board is
stored as a separate YAML file.

# Inspect the cells

You can list the cells of a capsule:

```{r}
capsule_cells("my_capsule")
```

You can inspect one cell:

```{r}
capsule_get_cell("my_capsule", "intro")
```

A cell contains the information needed to display one pedagogical unit on the
board: its title, section, position, text, code, unlocking question and links to
other cells.

# Understand `serious.yml`

The `serious.yml` file stores the general information about the capsule.

A simplified example looks like this:

```yaml
id: my_capsule
title: "My first SeRiouS capsule"
subtitle: "A shareable SeRiouS capsule folder"
method: "Introduction R"
description: "This capsule demonstrates the basic SeRiouS workflow."
language: en
type: serious_capsule
version: 0.2.0
start_cell: intro

packages:
  - datasets
  - graphics
  - stats

sections:
  - id: data
    label: "Data"
    color: "#E3F2FD"
    border: "#1565C0"
  - id: summary
    label: "Summary"
    color: "#E8F5E9"
    border: "#2E7D32"
  - id: visual
    label: "Visualisation"
    color: "#FFF3E0"
    border: "#EF6C00"
  - id: conclusion
    label: "Conclusion"
    color: "#F3E5F5"
    border: "#6A1B9A"
```

The `sections` field controls the legend and the colours used on the board.
The `start_cell` field indicates which cell is unlocked first.

# Understand a cell file

A cell is stored in `cells/` as a YAML file.

For example, a cell may contain:

```yaml
id: intro
title: Welcome
section: data
x: 0
y: 0
objective: Start the capsule.
text: Welcome to this SeRiouS capsule.
code: |
  data(iris)
  head(iris)
outputs:
  - console
expected_output: The first rows of the iris data set.
question: Which dataset is used?
expected_answer: iris
next_cells:
  - summary
```

This cell:

- appears in the `data` section;
- is placed at coordinates `x = 0`, `y = 0` on the board;
- displays an objective and explanatory text;
- runs R code;
- asks a question;
- links to the next cell, `summary`.

# Add a cell

You can add a new cell with `capsule_add_cell()`.

```{r}
capsule_add_cell(
  path = "my_capsule",
  id = "new_cell",
  title = "A new cell",
  section = "summary",
  x = 1000,
  y = 0,
  objective = "Introduce a new concept.",
  text = "This cell explains a new idea before running code.",
  code = "1 + 1",
  outputs = "console",
  expected_output = "The result is 2.",
  question = "What is 1 + 1?",
  expected_answer = "2"
)
```

This creates a new YAML file in the `cells/` folder.

# Connect cells

The learning path is defined by links between cells.

To connect an existing cell to the new cell:

```{r}
capsule_connect_cells(
  path = "my_capsule",
  from = "conclusion",
  to = "new_cell"
)
```

This means that `new_cell` becomes reachable after `conclusion`.

You can also remove a link:

```{r}
capsule_disconnect_cells(
  path = "my_capsule",
  from = "conclusion",
  to = "new_cell"
)
```

# Move a cell on the board

Cells have coordinates on the board.

To move a cell:

```{r}
capsule_move_cell(
  path = "my_capsule",
  id = "new_cell",
  x = 1250,
  y = 0
)
```

This changes the visual position of the cell in the board layout.

# Update an unlocking question

A cell can be associated with a question and an expected answer.

```{r}
capsule_set_cell_unlock(
  path = "my_capsule",
  id = "new_cell",
  question = "What is the result?",
  expected_answer = "2"
)
```

The answer is used by the Shiny application to unlock the next part of the
learning path.

# Check the capsule

Before sharing a capsule, it is recommended to check it.

From inside the capsule folder:

```{r}
check_capsule_dir(".")
```

Or from the parent folder:

```{r}
check_capsule_dir("my_capsule")
```

This function checks that:

```text
serious.yml exists
cells/ exists and contains valid cell files
the capsule can be loaded
the capsule can be validated
referenced PDF resources exist
the main folder structure is coherent
```

# Run the capsule

The main function for launching a capsule is `run_capsule()`.

To run a capsule from the parent folder:

```{r}
run_capsule("my_capsule")
```

To run it from inside the capsule folder:

```{r}
run_capsule(".")
```

This launches the Shiny interface.

The board shows the pedagogical cells. At the beginning, only the first cell is
unlocked. Learners progress by answering questions, running code, reading
outputs and moving from one cell to the next.

# Run an internal capsule

SeRiouS also includes internal demonstration capsules.

```{r}
available_capsules()
```

You can run an internal capsule directly:

```{r}
run_capsule("demo_iris")
```

or:

```{r}
run_capsule("demo_pca")
```

You can also retrieve a capsule as an R object:

```{r}
capsule <- get_capsule("demo_iris")
```

and run it:

```{r}
run_capsule(capsule)
```

# Share the capsule

To share the capsule, zip the folder:

```text
my_capsule/
```

The recipient only needs to install SeRiouS and run:

```{r}
library(SeRiouS)

run_capsule("my_capsule")
```

This makes it possible to share pedagogical material without creating a full R
package.

# Summary of the folder-based workflow

The typical workflow for a capsule author is:

```{r}
library(SeRiouS)

create_capsule_skeleton("my_capsule")

capsule_cells("my_capsule")
capsule_get_cell("my_capsule", "intro")

capsule_add_cell(
  path = "my_capsule",
  id = "new_cell",
  title = "A new cell",
  section = "summary",
  x = 1000,
  y = 0,
  text = "This is a new cell.",
  code = "1 + 1",
  question = "What is 1 + 1?",
  expected_answer = "2"
)

capsule_connect_cells(
  path = "my_capsule",
  from = "conclusion",
  to = "new_cell"
)

check_capsule_dir("my_capsule")
run_capsule("my_capsule")
```

The files to edit are:

```text
my_capsule/serious.yml
my_capsule/cells/*.yml
```

# Advanced: build a capsule directly in R

For package developers, or for users who want to generate capsules
programmatically, SeRiouS still provides a lower-level R API.

A pedagogical cell can be defined as a step:

```{r}
step1 <- make_step(
  id = "intro",
  title = "Inspect the data",
  section = "data",
  objective = "Display the first rows of the iris data set.",
  text = "We start by inspecting the iris data set.",
  code = "head(iris)",
  expected_output = "The first rows of the iris data set.",
  outputs = "console",
  question = "What is the name of the data set?",
  expected_answer = "iris"
)

step2 <- make_step(
  id = "summary",
  title = "Summarize the data",
  section = "summary",
  objective = "Compute a summary of the iris data set.",
  text = "We now compute a statistical summary.",
  code = "summary(iris)",
  expected_output = "A statistical summary of the variables.",
  outputs = "console"
)
```

Then the capsule can be built:

```{r}
capsule <- build_linear_capsule(
  id = "my_capsule",
  title = "My first SeRiouS capsule",
  method = "demo",
  steps = list(step1, step2),
  start_step = "intro",
  data = list(iris = iris)
)
```

The capsule can be validated and run:

```{r}
validate_capsule(capsule)
run_capsule(capsule)
```

This programmatic API is useful when capsules are generated from scripts or
included inside an R package. For most users, the recommended workflow is the
folder-based workflow with `serious.yml` and `cells/*.yml`.

# Conceptual model

The cell-based model can be summarised as follows:

```text
capsule = metadata + board

board = cells + links + sections

cell = pedagogical content + code + output + unlocking rule

link = connection from one cell to another
```

This model separates the pedagogical content from the running engine. It makes
capsules easier to inspect, edit, validate and share.
